Limits on constraining galaxy profiles with current and future strong gravitational lensing

### 2013-02-14 limits/ms.tex
More in the intro.

\tableofcontents

@@

If one assumes knowledge of the Hubble constant $H_0$ and time delays have been
observed for a lens. In principle, this determines that mass, but in practice
this can only place strong constraints on the mass distribution, as it does not
solve the problem that the exact distribution is highly degenerate. One can
choose to postpone this problem and simply fit models to the given data.  The
earliest examples were singular isothermal sphere models (SIS) but have since
evolved to multi-component models include separate models for the stars and
dark matter. Other methods that do not ignore this problem use genetic
algorithms or linear programming to explore the large space of solutions.

@@

\subsection{Future} %-----------------------------------------------------------
    \begin{itemize}
    \item New satellites will find many new lenses.
    \item LSST, DES, PanSTARRS?, SKA
    \item What is the most important data to have?
    \end{itemize}

\subsection{Present} %----------------------------------------------------------
    \begin{itemize}
    \item SLACS, CASTLES, CosmoGrail, CLASS, CFHTLS, SNAP?
    \end{itemize}

\subsection{Why do we need accurate galaxy models?} %---------------------------
    \begin{itemize}
    \item Galaxy formation
    \item Cusp/Core problem
    \item Universal halos?
    \item Adiabatic contraction
    \item DM fraction
    \end{itemize}

\subsection{Past work on constraining profiles.} %------------------------------
    \begin{itemize}
    \item Much of the early literature used simple SIS and SIE parametric models. Understandable.
    \item Bayesian model fitting
    \item Free-form modeling
    \item Keeton has a good review.
    \end{itemize}

\subsection{What are the problems?} %------------------------------
    \begin{itemize}
    \item Very little data per lens.
    \item Several well-known degeneracies.
    \item Velocity dispersions: aperture too small
    \end{itemize}

\subsection{Goals of the paper} %------------------------------
    \begin{itemize}
    \item Perform a parameter study over lens morphologies and quantify what they can tell us about profiles.
    \end{itemize}

>>>>

%\tableofcontents

@@

If one assumes knowledge of the Hubble constant $H_0$ and time delays have been
observed for a lens. In principle, this determines that mass, but in practice
this can only place strong constraints on the mass distribution, as it does not
solve the problem that the exact distribution is highly degenerate. One can
choose to postpone this problem and simply fit models to the given data.  The
earliest examples were singular isothermal sphere (SIS) or elliptical (SIE)
models but have since evolved to multi-component models include separate models
for the stars and dark matter. Other methods that do not ignore this problem
use genetic algorithms or linear programming to explore the large space of
solutions.

@@

Despite these difficulties lensing is a direct probe of the gravitating
mass in a halo and there are a number of open problems in galaxy formation
that lensing can help address. For instance, the debate remains as to what
form the inner slope of the density profile takes. Numerical simulations of
dark matter suggest a cusp, while observations from lensing and stellar dynamics
suggest cores. This may, however, be only a matter of scale, as both theory
and observations agree that clusters have cuspy profiles.

Lensing may also be useful in measuring possible adiabatic contraction of the 
dark matter halo due to the central stars in a galaxy. For this, one needs
high resolution mass models of the stellar profile combined with accurate models
of the total mass. Indeed, with such models, one may also improve the statistics
of the dark matter fraction in a galaxies over many different mass scales.

These open problems suggest what kind of lensing data will be required in the
future.  Future surveys will provide many lenses but time delays will require a
long follow up. Stellar mass maps, however, take a known, fixed observation
time and can be performed for many lenses. It will be important not only to
measure the stellar velocity dispersion at an arbitrary radius, but to ensure
that the measurement is taken at a radius different to the Einstein radius.
For the radio surveys like SKA it will be important to detect the central image
of the lens. This can provide mass information all the way to the center of the
galaxy.

In this paper we perform a parameter study using four mock galaxies that differ
in their dark matter and stellar profiles. We consider double and quad lensing
morphologies, time delays, the presence of a central image, red shift contrast
between multiple sources, and image configuration.  We also consider the
effects of adding stellar mass models as lower bound constraints on the mass
distribution.  The modeling and analysis is performed using a new free form
lens modeling framework called \Glass.

We first review the basic lensing theory we will need and then describe how
\Glass is used to generate lensing models. The mock data and how it was
generated is described in \secref{}. In section{} we discuss the lensing
configurations in the parameter study and present some example output. Our
finals results are presented in \secref{}.

%\subsection{Future} %-----------------------------------------------------------
    %\begin{itemize}
    %\item New satellites will find many new lenses.
    %\item LSST, DES, PanSTARRS?, SKA
    %\item What is the most important data to have?
    %\end{itemize}

%\subsection{Present} %----------------------------------------------------------
%    \begin{itemize}
%    \item SLACS, CASTLES, CosmoGrail, CLASS, CFHTLS, SNAP?
%    \end{itemize}

%\subsection{Why do we need accurate galaxy models?} %---------------------------
    %\begin{itemize}
    %\item Galaxy formation
    %\item Cusp/Core problem
    %\item Universal halos?
    %\item Adiabatic contraction
    %\item DM fraction
    %\end{itemize}

%\subsection{Past work on constraining profiles.} %------------------------------
    %\begin{itemize}
    %\item Much of the early literature used simple SIS and SIE parametric models. Understandable.
    %\item Bayesian model fitting
    %\item Free-form modeling
    %\item Keeton has a good review.
    %\end{itemize}

%\subsection{What are the problems?} %------------------------------
    %\begin{itemize}
    %\item Very little data per lens.
    %\item Several well-known degeneracies.
    %\item Velocity dispersions: aperture too small
    %\end{itemize}

%\subsection{Goals of the paper} %------------------------------
%    \begin{itemize}
%    \item Perform a parameter study over lens morphologies and quantify what they can tell us about profiles.
%    \end{itemize}
